---
title: "Guidelines to a reproducible analysis"
author: "Gerard Romero Sola"
date: "2023-03-08"
output: html_document
---

First, I created a github repository online, since the following command will create a local one from the one created via R. To create the repository I followed this guidelines:

* Repository template: No template.

* Repository name: myrepo or whatever you wish to name your new project. Approach this similar to a variable name, in code: descriptive but brief, no whitespace. Letters, digits, -, ., or _ are allowed.

**IMPORTANT:**
For further steps, since the repository is treated like a package, it's important not to use the - or _. Instead, follow this rules:
… contain only ASCII letters, numbers, and ‘.’
… have at least two characters
… start with a letter (not a number)
… not end with ‘.’

* Description: “Analysis of the stuff” or any short description of the project. Write this for humans.

* Public.

* Initialize this repository with: Add a README file.

#### Downloading the tools
The required package to conduct this process is `rrtools`. To install it:
```{r}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("benmarwick/rrtools")
```

### New RStudio Project via git clone
[Link](https://happygitwithr.com/new-github-first.html)
I directly tried to use the `usethis::create_from_github()` command, but the credentials were wrong. To check them:
```{r}
gitcreds::gitcreds_set()
```

First, I checked the credentials were okay, and they were not, after selecting 3 I realised the password was wrong. So I selected 2 and replace it with the proper one.

So, now that everything is correct:
```{r}
usethis::create_from_github(
	"https://github.com/gerirosoupf/rrtools.trial.git",
	destdir = "~/Desktop/Analysis/"
)
```

There’s a big advantage to the “GitHub first, then RStudio” workflow: the remote GitHub repo is configured as the `origin` remote for your local repo and your local `main` branch is now tracking the main on GitHub. This is a technical but important point about Git. The practical implication is that you are now set up to push and pull. No need to fanny around setting up Git remotes and tracking branches on the command line.

### rrtools::use_compendium("pkgname")

If you started with a new project on GitHub first, run rrtools::use_compendium(), if you started with a new project in RStudio first, run rrtools::use_compendium("pkgname")
```{r}
rrtools::use_compendium()
```

This uses usethis::create_package() to create a basic R package in the pkgname directory, and then, if you’re using RStudio, opens the project. If you’re not using RStudio, it sets the working directory to the pkgname directory.
we need to:
* edit the DESCRIPTION file (located in your pkgname directory) to include accurate metadata, e.g. your ORCID and email address

* periodically update the Imports: section of the DESCRIPTION file with the names of packages used in the code we write in the qmd document(s) by running rrtools::add_dependencies_to_description(). If I do it, it takes into account the packages used in this RMarkdown document.
```{r}
rrtools::add_dependencies_to_description()
```

### usethis::use_mit_license(copyright_holder = "My Name")

```{r}
usethis::use_mit_license(copyright_holder = "Gerard Romero")
```

* this adds a reference to the MIT license in the DESCRIPTION file and generates a LICENSE file listing the name provided as the copyright holder

* to use a different license, replace this line with any of the licenses mentioned here: ?usethis::use_mit_license()

### rrtools::use_readme_rmd()

```{r}
rrtools::use_readme_rmd()
```
this generates README.Rmd and renders it to README.md, ready to display on GitHub. It contains:

* a template citation to show others how to cite your project. Edit this to include the correct title and DOI.

* license information for the text, figures, code and data in your compendium

this also adds two other markdown files: a code of conduct for users CONDUCT.md, and basic instructions for people who want to contribute to your project CONTRIBUTING.md, including for first-timers to git and GitHub.

* this adds a .binder/Dockerfile that makes Binder work, if your compendium is hosted online. Currently configured for GitHub, but easily adapted for elsewhere (e.g. Zenodo, Figshare, Dataverse, etc.)

* render this document after each change to refresh README.md, which is the file that GitHub displays on the repository home page (basically, change the Rmd and then knit)

### rrtools::use_analysis()

```{r}
rrtools::use_analysis()
```

* this function has three location = options: top_level to create a top-level analysis/ directory, inst to create an inst/ directory (so that all the sub-directories are available after the package is installed), and vignettes to create a vignettes/ directory (and automatically update the DESCRIPTION). The default is a top-level analysis/.

* for each option, the contents of the sub-directories are the same, with the following (using the default analysis/ for example):

* the paper.qmd is ready to write in and render with Quarto. It includes:

- a YAML header that identifies the references.bib file and the supplied csl file (to style the reference list)

- a colophon that adds some git commit details to the end of the document. This means that the output file (HTML/PDF/Word) is always traceable to a specific state of the code.

* the references.bib file has just one item to demonstrate the format. It is ready to insert more reference details.

* you can replace the supplied csl file with a different citation style from https://github.com/citation-style-language/
we recommend using the RStudio 2022.07 or higher to efficiently insert citations from your Zotero library while writing in an qmd file (see here for detailed setup and use information to connect your RStudio to your Zotero)
remember that the Imports: field in the DESCRIPTION file must include the names of all packages used in analysis documents (e.g. paper.qmd). We have a helper function rrtools::add_dependencies_to_description() that will scan the qmd file, identify libraries used in there, and add them to the DESCRIPTION file.
this function has an data_in_git = argument, which is TRUE by default. If set to FALSE you will exclude files in the data/ directory from being tracked by git and prevent them from appearing on GitHub. You should set data_in_git = FALSE if your data files are large (>100 mb is the limit for GitHub) or you do not want to make the data files publicly accessible on GitHub.
To load your custom code in the paper.qmd, you have a few options. You can write all your R code in chunks in the qmd, that’s the simplest method. Or you can write R code in script files in /R, and include devtools::load_all(".") at the top of your paper.qmd. Or you can write functions in /R and use library(pkgname) at the top of your paper.qmd, or omit library and preface each function call with pkgname::. Up to you to choose whatever seems most natural to you.



